I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-04-27 lipics-master/experiments.tex

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes from layers $1$ and $2$. We test using compensation, compared with a random choice, and also choice based on strength in \cite{9211410}, defined as $strength(c) = (UB(n)-LB(n))\cdot W_{cd}$. It is however limited to chosing ReLU in the previous layer 2. This choice is fine for few nodes (10,20), as there is a high corelation between $comp(c)$ being high and $W_{cd}$ being high. However, as soon as many nodes can be chosen in $Z$, compensate becomes much more accurate. Even random becomes more accurate as the number of nodes grows, as \cite{9211410} only has access to nodes $c$ in layer 2, and not to nodes $b$ in layer 1.
For comparison, compensate is the most accurate, from 30 nodes, experimentally validating that compensating strength is the important factor to consider.

@@

CMP ranks highly in terms of accuracy, on the moderatly large network (but challenging verification-agnostic) DNNs we experimented with. Even if the CMP complexity is not subject to combinatorial explosion, its runtime, particularly for larger DNNs, can be relatively slow, as CMP needs to consider every node one by one. Tackling much larger networks may thus need different implementation than CMP, while still benefitting from the novel compensation notion we uncovered and validated thanks to CMP. Numerous avenues exist:

>>>>

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes from layers $1$ and $2$. We test using compensation, compared with a random choice, and also choice based on strength in \cite{9211410}, defined as $strength(c) = (UB(n)-LB(n))\cdot W_{cd}$. It is however limited to choosing ReLU in the previous layer 2. This choice is fine for few nodes (10,20), as there is a high correlation between $comp(c)$ being high and $W_{cd}$ being high. However, as soon as many nodes can be chosen in $Z$, compensate becomes much more accurate. Even random becomes more accurate as the number of nodes grows, as \cite{9211410} only has access to nodes $c$ in layer 2, and not to nodes $b$ in layer 1.
For comparison, compensate is the most accurate, from 30 nodes, experimentally validating that compensating strength is the important factor to consider.

@@

CMP ranks highly in terms of accuracy, on the moderately large network (but challenging verification-agnostic) DNNs we experimented with. Even if the CMP complexity is not subject to combinatorial explosion, its runtime, particularly for larger DNNs, can be relatively slow, as CMP needs to consider every node one by one. Tackling much larger networks may thus need different implementation than CMP, while still benefitting from the novel compensation notion we uncovered and validated thanks to CMP. Numerous avenues exist:

### 2024-04-27 lipics-master/relatedwork.tex

We compare {\CMP} with major verification tools for DNNs to clarify our methodology and its distinction from the existing state-of-the-art. CMP scales while preserving good accuracy, through targeting a limited number of binary variables based on compensating strength, stricking a good balance between exact encoding of a DNN using MILP~\cite{MILP} (too slow) and LP relaxation (too inaccurate). MIPplanet~\cite{MIPplanet} opts for a different selection of binary variables, and execute one large MILP instance instead of CMP's many small instances, which significantly reduce the number of binary variables necessary for each instance. In \cite{9211410}, small instances are also considered, however with a straightfoward choice of binary nodes based on the weight of outgoing edge, therefore restricted to nodes in the previous layer, which is significantly less accurate than using compensating strength (see Table \ref{tab:example1}).

>>>>

We compare {\CMP} with major verification tools for DNNs to clarify our methodology and its distinction from the existing state-of-the-art. CMP scales while preserving good accuracy, through targeting a limited number of binary variables based on compensating strength, stricking a good balance between exact encoding of a DNN using MILP~\cite{MILP} (too slow) and LP relaxation (too inaccurate). MIPplanet~\cite{MIPplanet} opts for a different selection of binary variables, and execute one large MILP instance instead of CMP's many small instances, which significantly reduce the number of binary variables necessary for each instance. In \cite{9211410}, small instances are also considered, however with a straightforward choice of binary nodes based on the weight of outgoing edge, therefore restricted to nodes in the previous layer, which is significantly less accurate than using compensating strength (see Table \ref{tab:example1}).
